Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective

### 2021-11-13 edubots_surveys.Rnw
Changing a word
Fixing one suggestion
Restructured the Discussion and Conclusion sections
I have collated the discussion and the conclusion sections into one section.  All the extra text I copied it at the bottom of the page in a section marked 'Extra'.  I will work again on it tomorrow.

The {\bf initial survey} was done in two separate Google forms, one for
university and other for tertiary non-university teachers. Questions and
responses were the same, except for the type of tertiary degrees that were
taught. In both cases, we used mainly mailing lists, as well as Telegram groups,
to spread the link to the form. The university form was spread in Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also Costa Rica and
Mexico. The tertiary non-university teachers are based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, post-pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.

@@

% ********************** Discussion section *******************
\section{Discussion}

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers;
apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the mind of the
surveyed.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few people actually
changed, with women vs. men being an exception, as well as those teaching social
sciences standing out among the rest. We do not have any kind of explanation for
this behaviour, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions that can arise from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? \textcolor{blue}{The survey 
only listed a few functionalities for the chatbots and only a few subjects selected 
the option "Others". This may suggest  that most teachers did not really understand 
at this point the actual potential of chatbots.} Another way to look about it is 
that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burn out to think about
innovation.

It is also likely that their opinions towards chatbots are informed by negative
experiences with customer support chatbots, which is what most people will have
experienced. This will probably require further study, but also will need to be
taken into account when introducing chatbot technologies in the classroom: 
\textcolor{blue}{it will have to be as simple, and as satisfactory, as possible, 
to (possibly) change this existing perception}.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. When looking at the plot containing the sector differentiation
it is interesting to observe how university teachers prefer in a higher
proportion to use them for grades and FAQs in comparison with Vocational High
School teachers. \textcolor{blue}{
This seems reasonable as the frequency of seeing the students
is lower at the University so, an automatic tool for grade notification for
continuous evaluation seems more helpful}. 
Furthermore, the high student/teacher
ratios that university teachers have to deal with could be another reason to
have a higher number of them willing to use chatbots for FAQs. This could
ameliorate the number of misunderstandings and email overload that so many
students might generate.

The results, when grouping by experience, confirm the previous observation that
experienced teachers are not afraid of changes and are willing to accept new
technological challenges, at least, in the same way that younger teachers
are. From this concrete plot it is interesting as well the fact that very few
young/inexperience teachers are willing to use chatbots for other uses. \textcolor{blue}{This
reinforces the hypothesis thrown at the beginning of this section that younger teachers have less experience to consider new ways of using chatbots in their daily work.}

As it appears in the responses presented above, teachers’ responses are
distributed between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone, while for the features: familiarity,
horizontally, sustainability and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the
majority of teachers seems to consider not valuable the presence of official
formation, pluggability, unidirectionality and synchrony in the interaction
media features of messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations and to avoid the
   overhead required to \textcolor{blue}{sign up new students every year}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}

An analysis of survey questions related to RQ4 seems to indicate that teachers
prefer to have interaction with their students in a chat "room"; a preference of
{\em horizontal} interaction might indicate the need to delegate or offload part
of the burden of answering, and attending, every single question posed by
students. In an horizontal setting, other students can answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
that way, activity in the chat tool, answering questions, regularity, can all be
observed and valued as part of the process of assessing achievement of learning
objectives by students.

The key findings of this study shed the light on educators’ preference to use
messaging applications that their institutions support.  Technology adoption
literature often \textcolor{blue}{focuses} on users’ perceptions of the technology’s usefulness and
ease of use as important prerequisites for successful adoption and
utilization. Nevertheless, in higher education, institutions’ role in
integrating these tools to their educational systems can improve the uptake of
these applications and shape the social and educational experiences of their
students. \textcolor{blue}{To accomplish this fruitful integration}, institutions have to ensure that
these messaging applications not only are GDPR compliant to keep students’ data
secure but also should provide IT support to all stakeholders who use the
applications.

Moreover, this study clarifies the differences between the teachers’
perspectives and/or intentions and that of their students when using the
messaging applications in higher education.  Teachers tend to adopt technologies
supported by their institutions.  This could be to ensure that their
universities can oversee their efforts in supporting their students during the
learning process.  Another reason could be their familiarity with the technology
provided by their institutions. On the other hand, students use the
non-institution messaging applications to form informal discussion groups with
their peers.  It is worth noting that peer support and collaboration is
inseparable from learning \cite{timmis2012constant} and correlates positively
with higher retention in higher education \cite{o2014front}.  Therefore, both
perspectives are complementary and play different roles in promoting the
learning process.

% Noorhan suggest that using some simple things could increase the perceived usefulness - JJ
As in many other methods of technology acceptance, answers to the survey suggest
that the introduction of simple, and institution-supported, instant messaging
and chatbot technologies would increase the perceived usefulness (which is one
of the key metrics in technology acceptance models). Since most institutions
already have some {\em virtual campus} or learning management system, adding
some easy automation, or connection to personal instant messaging tools, could
really help onboard the learning community on these new technologies.

% What kind of chatbots are preferred? (suggested by Fran)
Once that initial introduction is done, teachers (and students as well) will
probably prefer the kind of bots that alleviate bureaucratic or repetitive
tasks, such as answering frequent questions or answers on class or assignment
schedules. These will help the introduction of more complex chatbots that will
affect more directly learning outcomes, such as chatbots that help students
integrate in the class, or are able to identify (and address) learning problems
in students or in groups of them. These as well should be accompanied by
analytics on student interaction, as well as possibly some natural language
processing (in vernacular language) that will help assess the general mood of
the class, and how different material (or external factors) affect it.

@@

>>>>

The {\bf initial survey} was done in two separate Google forms, one for
university and other for tertiary non-university teachers. Questions and
responses were the same, except for the type of tertiary degrees that
\textcolor{blue}{were considered}. In both cases, we used mainly mailing lists, as well as Telegram groups,
to spread the link to the form. The university form was spread in Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also Costa Rica and
Mexico. The tertiary non-university teachers are based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, post-pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.

@@

% ********************** Discussion section *******************
\section{Discussion and Conclusion}

This study sheds the light on educators' growing usage of messaging applications in classrooms even before COVID19.  Recent research \cite{panah2020study} 
examining the use of messaging applications in higher education claimed that WhatsApp is widely used by educators. 
However, in this study, although 44% of educators reported using WhatsApp in their classrooms, more than half of the survey responses (nearly 56%)
claim using technologies supported by their institutions. One possible explanation for this finding could be that educators wanted to ensure that 
their institutions can oversee their efforts in supporting their students during the learning process.  Another reason could be attributed to 
institutions mandating the use of specific technologies that they support.  Hence, this finding probes more queries about the actual reasons 
contributing to this shift in technology preference among educators and its implications on the students/educators interaction and learning. 
In addition, as recent research examining students' usage of messaging applications with their peers in higher education \cite{abbas2021onlinechat}
argue that students prefer to use the non-institution messaging applications to form informal discussion groups with their peers, further 
research should aim to investigate this wide spectrum of usage of these applications and their impact on the learning process.

The question examining the use of messaging applications during the pandemic (COVID19) reflects the big challenge of having to 
change habits during a major crisis. The survey responses reveal some gender differences in this respect as 31% of female educators 
reported moving to using messaging applications to cope with the new virtual mode of learning during the pandemic compared to only 17% of male educators.
Moreover, the survey responses to the question that lists the different functionalities that a chatbot can be used for indicate that nearly 60%
of educators favoured using chatbots in grading and 47.5% in answering students' FAQs.

The survey results, when grouping by experience, confirm the previous observation that
experienced teachers are willing to accept new technological challenges, at least, similar to younger teachers. Younger educators 
are also willing to use chatbots for other uses.

Therefore, as in many other methods of technology acceptance, answers to the survey suggest
that the introduction of simple, and institution-supported, instant messaging
and chatbot technologies would increase the perceived usefulness (which is one
of the key metrics in technology acceptance models). Since most institutions
already have some {\em virtual campus} or learning management system, adding
some easy automation, or connection to personal instant messaging tools, could
really help on board the learning community on these new technologies.

% What kind of chatbots are preferred? (suggested by Fran)
Once that initial introduction is done, teachers (and students as well) will
probably prefer the kind of chatbots that alleviate bureaucratic or repetitive
tasks, such as answering frequent questions or answers on class or assignment
schedules. These will help the introduction of more complex chatbots that will
affect more directly learning outcomes, such as chatbots that help students
integrate in the class, or are able to identify (and address) learning problems
in students or in groups of them. These as well should be accompanied by
analytics on student interaction, as well as possibly some natural language
processing (in vernacular language) that will help assess the general mood of
the class, and how different material (or external factors) affect it.

@@

%***********************************Extra************************************************

Technology adoption literature often \textcolor{blue}{focuses} on users’ perceptions of the technology’s usefulness and
ease of use as important prerequisites for successful adoption and
utilization. Nevertheless, in higher education, institutions’ role in
integrating these tools to their educational systems can improve the uptake of
these applications and shape the social and educational experiences of their
students. \textcolor{blue}{To accomplish this fruitful integration}, institutions have to ensure that
these messaging applications not only are GDPR compliant to keep students’ data
secure but also should provide IT support to all stakeholders who use the
applications.
As it appears in the responses presented above, teachers’ responses are
distributed between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone, while for the features: familiarity,
horizontally, sustainability and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the
majority of teachers seems to consider not valuable the presence of official
formation, pluggability, unidirectionality and synchrony in the interaction
media features of messaging apps.
From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations and to avoid the
   overhead required to \textcolor{blue}{sign up new students every year}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}
An analysis of survey questions related to RQ4 seems to indicate that teachers
prefer to have interaction with their students in a chat "room"; a preference of
{\em horizontal} interaction might indicate the need to delegate or offload part
of the burden of answering, and attending, every single question posed by
students. In an horizontal setting, other students can answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
that way, activity in the chat tool, answering questions, regularity, can all be
observed and valued as part of the process of assessing achievement of learning
objectives by students.

% Noorhan suggest that using some simple things could increase the perceived usefulness - JJ
